---
title: "Statistical assignment 4"
author: "Connor Kowalewski"
date: "26/02/2020"
output: github_document
---

```{r setup, include=FALSE}
# Please note these options.
# This tells R Markdown that we want to show code in the output document.
knitr::opts_chunk$set(echo = TRUE)
# Switching off messages in the output document. 
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
# Switching on caching to make things faster (don't commit cache files on Github).
knitr::opts_chunk$set(cache = TRUE)
```

In this assignment you will need to reproduce 5 ggplot graphs. I supply graphs as images; you need to write the ggplot2 code to reproduce them and knit and submit a Markdown document with the reproduced graphs (as well as your .Rmd file).

First we will need to open and recode the data. I supply the code for this; you only need to change the file paths.

    ```{r}
library(tidyverse)
Data8 <- read_tsv("C:/Users/conno/Downloads/DA III Data/UKDA-6614-tab/tab/ukhls_w8/h_indresp.tab")
Data8 <- Data8 %>%
        select(pidp, h_age_dv, h_payn_dv, h_gor_dv)
Stable <- read_tsv("C://Users/conno/Downloads/DA III Data/UKDA-6614-tab/tab/ukhls_wx/xwavedat.tab")
Stable <- Stable %>%
        select(pidp, sex_dv, ukborn, plbornc)
Data <- Data8 %>% left_join(Stable, "pidp")
rm(Data8, Stable)
Data <- Data %>%
        mutate(sex_dv = ifelse(sex_dv == 1, "male",
                           ifelse(sex_dv == 2, "female", NA))) %>%
        mutate(h_payn_dv = ifelse(h_payn_dv < 0, NA, h_payn_dv)) %>%
        mutate(h_gor_dv = recode(h_gor_dv,
                         `-9` = NA_character_,
                         `1` = "North East",
                         `2` = "North West",
                         `3` = "Yorkshire",
                         `4` = "East Midlands",
                         `5` = "West Midlands",
                         `6` = "East of England",
                         `7` = "London",
                         `8` = "South East",
                         `9` = "South West",
                         `10` = "Wales",
                         `11` = "Scotland",
                         `12` = "Northern Ireland")) %>%
        mutate(placeBorn = case_when(
                ukborn  == -9 ~ NA_character_,
                ukborn < 5 ~ "UK",
                plbornc == 5 ~ "Ireland",
                plbornc == 18 ~ "India",
                plbornc == 19 ~ "Pakistan",
                plbornc == 20 ~ "Bangladesh",
                plbornc == 10 ~ "Poland",
                plbornc == 27 ~ "Jamaica",
                plbornc == 24 ~ "Nigeria",
                TRUE ~ "other")
        )
    ```

Reproduce the following graphs as close as you can. For each graph, write two sentences (not more!) describing its main message.  

1. Univariate distribution (20 points)
    ```{r}
Data %>%
     ggplot(aes(x = h_payn_dv)) +
     geom_freqpoly() +
     labs(x = "Net monthly pay",
          y = "Number of respondents")
    ```

Interpretation: We can see from this graph that the majority of respondents had a net monthly pay of between 0 and 3000. It is also slightly skewed to the right. 

2. Line chart (20 points) The lines show the non-parametric association between age and monthly earnings for men and women.
    ```{r}
    Data %>%
        ggplot(aes(x = h_age_dv, y = h_payn_dv, linetype =       sex_dv)) + 
        geom_smooth(colour = "black") + coord_cartesian(xlim = c(15, 65), ylim = c(0, 2250)) +
        scale_y_continuous(breaks = seq(500, 2000, by = 500)) +
        labs(x = "Age",
             y = "Monthly earnings", 
        linetype = "Sex")
    ```

Interpretation: This chart suggests that males on average have a higher monthly earning than females. It also demonstrates that monthly earnings are likely to peak for individuals during their 40s. 

3. Faceted bar chart (20 points)
    ```{r}
Data %>%
     filter(!is.na(h_payn_dv), !is.na(placeBorn), !is.na(sex_dv)) %>%
     group_by(placeBorn, sex_dv) %>%
     summarise(meanPAY = mean(h_payn_dv, na.rm = TRUE)) %>%
     ggplot(aes(x = sex_dv, y = meanPAY)) +
     geom_bar(stat = "identity") +
     facet_wrap(~ placeBorn, nrow = 3) +
     labs(x = "Sex", y = "Median monthly net pay")
    ```

Interpretation: These charts again suggest that regardless of place of birth, males have a higher median monthly net pay than females. They also demonstrate that Ireland has the highest median monthly net pay whilst Pakistan have the lowest. 

4. Heat map (20 points)
    ```{r}
    Data %>% 
        filter(!is.na(placeBorn), !is.na(h_gor_dv)) %>%
        group_by(placeBorn, h_gor_dv) %>%
        summarise(meanAGE = mean(h_age_dv, na.rm = TRUE)) %>%
        ggplot(aes(x = h_gor_dv, y = placeBorn)) +    
        geom_tile(aes(fill = meanAGE)) +
        theme_classic() + 
        theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
        coord_equal() +
        labs(x = "Region", y = "Country of birth", fill = "Mean age")
    ```

Interpretation: This heat map shows that the mean age is consistent for UK born individuals across all regions. The youngest mean age is Nigerians living in Scotland whilst the oldest mean age is Jamaicans living in Scotland.

5. Population pyramid (20 points)
    ```{r}
ggplot(data = Data, aes(x = h_age_dv, fill = sex_dv)) + 
geom_bar(data = dplyr::filter(Data, sex_dv == "female")) + 
geom_bar(data = dplyr::filter(Data, sex_dv == "male"), aes(y =    ..count.. * (-1))) + 
scale_fill_brewer(palette = "Set1") +
coord_flip() +
labs(x = "Age", y = "n", fill = "Sex")
    ```

Interpretation: This pyramid suggests that there were more female respondents during this wave. It also shows that the distribution of age between the sexes is consistent. 